---
title: "p8105_hw3_mp3653"
author: "Matthew Perrotta"
date: "October 10, 2018"
output: github_document
---

### Load packages
```{r}
library(tidyverse)
library(knitr)
library(patchwork)
```

# Problem 1
### Import data
```{r Import BRFSS_smart2010 data}
brfss = p8105.datasets::brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  filter(topic == 'Overall Health') %>% 
  filter(response %in% c('Excellent', 'Very good', 'Good', 'Fair', 'Poor')) %>% 
  mutate(response = factor(response, levels = c('Excellent', 'Very good', 'Good', 'Fair', 'Poor'))) %>% 
  rename(state = locationabbr, county = locationdesc)
```

### Questions
```{r Number of states with 7 observations}
brfss %>% 
  filter(year == 2002) %>% 
  group_by(state, county) %>% 
  summarize(n = n()) %>% 
  count(state) %>% 
  filter(nn == 7)
  
```

Create a spaghetti plot

```{r Make a 'spaghetti' plot}
brfss %>%
  distinct(year, state, county) %>% 
  group_by(year, state) %>% 
  summarize(n = n()) %>% 
  ggplot(aes(x = year, y = n, color = state)) +
  geom_line()
  
  
```

Create a table displaying the mean of excellent responses across NY counties for the years 2002, 2006, and 2010
```{r create table}
brfss %>% 
  filter(response == 'Excellent',
         state == 'NY',
         year %in% c(2002, 2006, 2010)) %>% 
  group_by(year) %>% 
  summarize(mean_excellent = mean(data_value/100),
            sd_excellent = sd(data_value/100)) %>% 
  knitr::kable()
```

Create plots displaying average response proportion for each state per year
```{r plots for average response proportion for each state per year}
brfss %>% 
  group_by(state, year, response) %>% 
  summarize(mean_values = mean(data_value/100)) %>% 
  ggplot(aes(x = year, y = mean_values, color = state)) +
  geom_line() +
  facet_grid(~response)
```

# Problem 2
### Import data
```{r Import instacart data}
instacart = p8105.datasets::instacart %>% 
  janitor::clean_names()
```

DESCRIBE DATA

Number of aisles
```{r number of aisles in instacart}
instacart %>% 
  distinct(aisle_id) %>% 
  nrow()
```
There are 134 aisles

Aisles with the most products ordered
```{r }
instacart %>% 
  group_by(aisle) %>% 
  count() %>% 
  arrange(desc(n))
```
The top three aisles with the most products ordered are fresh vegatables, fresh fruits, and pacjaged vegetables fruits.

Plot of number of items ordered in each aisle
```{r Plot of number of items ordered in each aisle}
instacart %>% 
  group_by(aisle, department) %>% 
  count() %>% 
  ggplot(aes(x = aisle, y = n)) +
  geom_point()
```

Table of most popular item in the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”
```{r}
instacart %>% 
  filter(aisle %in% c('baking ingredients', 'dog food care', 'packaged vegetables fruits')) %>% 
  group_by(aisle) %>% 
  count(product_name) %>% 
  filter(min_rank(desc(n)) < 2) %>% 
  rename(number_of_orders = n)
  knitr::kable()
```

Table of mean ordering hour each day of the week for Pink Lady Apples and Coffee Ice Cream
```{r Mean ordering hour for each day of the week}
instacart %>% 
  filter(product_name %in% c('Pink Lady Apples', 'Coffee Ice Cream')) %>% 
  group_by(order_dow, product_name) %>% 
  summarise(mean_order_hour = round(mean(order_hour_of_day))) %>% 
  spread(key = order_dow, value = mean_order_hour) %>% 
  rename(Monday = '0', Tuesday = '1', Wednesday = '2', Thursday = '3', Friday = '4', Saturday = '5', Sunday = '6') %>% #assuming day 0 is monday
  knitr::kable()
```

#Problem 3
### Data import
```{r Import NOAA data for NY}
ny_noaa = p8105.datasets::ny_noaa %>% 
  janitor::clean_names() %>% 
  separate(date, into = c('year', 'month', 'day'), sep = '-') %>% 
  mutate(tmax = as.numeric(tmax),
         tmin = as.numeric(tmin)) %>% 
  mutate(prcp = prcp/10, 
         tmax = tmax/10, 
         tmin = tmin/10)
```

The `ny_noaa` dataset is `r nrow(ny_noaa)` rows by `r ncol(ny_noaa)` columns and is in long format. ....

Plot of average max temp in the months of January and July for each station across years
```{r plot of max temps}
ny_noaa %>% 
  filter(month %in% c('01', '07')) %>% 
  group_by(id, year, month) %>% 
  summarise(mean_tmax = mean(tmax)) %>% 
  ggplot(aes(x = year, y = mean_tmax)) +
  geom_point() +
  facet_grid('month') +
  theme(legend.position = 'none')
```


